### 2014-09-30 paper/inputs/tab_mc_overview.tex
add missing file

%!TEX root = ../egpaper_for_review.tex
\begin{table}
\begin{center}
    \begin{tiny}
    \begin{tabular}{ p{1cm} p{1cm} p{1.5cm} p{2.5cm} }
    \hline
    \hline
    Method &    & Limitations &   \\ \hline
    Multicut \cite{andres_2011_iccv,kappes_2011_emmcvpr} & 
        (I)LP & 
        not scalable, slow &
        Cutting plane method.
        Global optimal but slow for huge problems. 
        However, applicable for medium size superpixel RAG. 
        \\ \hline
    Expand And Explorer \cite{bagon_2011_arxiv} & 
        movemaker & 
        gets stuck early &
        $\alpha$-Expansion modification. Applicable to
        large scale problems but gets stuck in local minima(???).
        \\ \hline
    Fast Planar CC \cite{yarkony_2012_eccv} & 
        dual-decompositon & 
        planar graphs only, 
        duality/integrality gab &
        Dual decompositon in planar subgraphs??
        Works on planar graphs only. Fast (is there a word for fast to medium fast) for
        small to medium size graphs.
        \\ \hline
    Cut Glue and Cut \cite{beier_2014_cvpr} & 
        move-making & 
        gets stuck in poor minima for non planar graphs &
        Greedy solver based on a series of 2 colorings.
        State of the art any time performance. Very
        fast for planar graphs. Gets stuck in local minima
        for non planar graphs.
        \\ \hline
    THIS WORK & 
        fusion move-making & 
        no natural stopping condition &
        Fusion of a proposal segmentation with
        the current best segmentation.
        State of the art any time performance for (???)
        \\ \hline
        \hline
    \end{tabular}
    \end{tiny}
\end{center}
\caption{
    Overview of correlation clustering / multicut solvers
}
\end{table}