### 2014-10-13 paper/tables/seg-3d.tex

\begin{table}[H]
\scriptsize
\centering
\caption{seg-3d (2 instances)}
\label{tab:smalltable-seg-3d}
\begin{tabular}{lrrrrrr}
\toprule
           algorithm &       runtime     &         value &         bound &           mem &     best &      opt   \\ \midrule 
             ogm-ICM & $      1959.44$ sec & $    618075.17$ & $-\infty$ & $         0.22$ GB & $       0$ & $       0$ \\ 
              ogm-KL & $      2290.46$ sec & $    441695.84$ & $-\infty$ & $         0.20$ GB & $       0$ & $       0$ \\ 
            ogm-LF-1 & $      1857.93$ sec & $    552690.10$ & $-\infty$ & $         0.27$ GB & $       0$ & $       0$ \\ 
\cmidrule{1-1} 
              MCR-CC & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.11$ GB & $       1$ & $       1$ \\ 
           MCR-CCFDB & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.05$ GB & $       1$ & $       1$ \\ 
       MCR-CCFDB-OWC & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.05$ GB & $       1$ & $       1$ \\ 
\cmidrule{1-1} 
     MCI-CCFDB-CCIFD & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.05$ GB & $       1$ & $       1$ \\ 
             MCI-CCI & $      1806.76$ sec & $    414124.76$ & $    413523.92$ & $         4.43$ GB & $       1$ & $       1$ \\ 
           MCI-CCIFD & $      4201.35$ sec & $    650888.58$ & $    401952.16$ & $         1.01$ GB & $       1$ & $       1$ \\ 
\cmidrule{1-1} 
            ogm-CGC* & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.03$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE* & $        27.97$ sec & $    501728.60$ & $-\infty$ & $         0.38$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE-CGCF* & $        40.77$ sec & $    471105.16$ & $-\infty$ & $         0.55$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-CGC* & $      1293.80$ sec & $    413546.15$ & $-\infty$ & $         0.72$ GB & $       1$ & $       0$ \\ 
ogm-mcfusion-HC-CGC-CGCF* & $      1291.94$ sec & $    413545.50$ & $-\infty$ & $         0.88$ GB & $       2$ & $       0$ \\ 
 ogm-mcfusion-HC-MC* & $      1171.51$ sec & $    413552.74$ & $-\infty$ & $         0.68$ GB & $       1$ & $       0$ \\ 
ogm-mcfusion-HC-MC-CGCF* & $      1172.41$ sec & $    413546.96$ & $-\infty$ & $         0.85$ GB & $       1$ & $       0$ \\ 
ogm-mcfusion-WS-BASE* & $         7.56$ sec & $    748369.02$ & $-\infty$ & $         0.19$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-BASE-CGCF* & $      6088.09$ sec & $    413807.74$ & $-\infty$ & $         0.40$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-CGC* & $      1838.10$ sec & $    413760.67$ & $-\infty$ & $         0.28$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-CGC-CGCF* & $      2484.01$ sec & $    413573.58$ & $-\infty$ & $         0.44$ GB & $       0$ & $       0$ \\ 
 ogm-mcfusion-WS-MC* & $     32840.49$ sec & $    427635.00$ & $-\infty$ & $        10.53$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-MC-CGCF* & $     34279.76$ sec & $    413814.27$ & $-\infty$ & $        10.69$ GB & $       1$ & $       0$ \\ 
\bottomrule
\end{tabular}
\end{table}