### 2014-10-13 paper/tables/image-seg.tex

\begin{table}[H]
\scriptsize
\centering
\caption{image-seg (100 instances)}
\label{tab:smalltable-image-seg}
\begin{tabular}{lrrrrrr}
\toprule
           algorithm &       runtime     &         value &         bound &           mem &     best &      opt   \\ \midrule 
     ogm-CGC-planar* & $         0.28$ sec & $      4445.22$ & $      4136.83$ & $         0.02$ GB & $      28$ & $       0$ \\ 
ogm-mcfusion-HC-BASE* & $         0.12$ sec & $      5462.60$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE-CGCF* & $         0.12$ sec & $      5085.80$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-CGC* & $         1.23$ sec & $      4444.56$ & $-\infty$ & $         0.01$ GB & $      40$ & $       0$ \\ 
ogm-mcfusion-HC-CGC-CGCF* & $         1.22$ sec & $      4444.75$ & $-\infty$ & $         0.02$ GB & $      40$ & $       0$ \\ 
 ogm-mcfusion-HC-MC* & $         5.12$ sec & $      4443.61$ & $-\infty$ & $         0.04$ GB & $      73$ & $       0$ \\ 
ogm-mcfusion-HC-MC-CGCF* & $         5.14$ sec & $      4443.61$ & $-\infty$ & $         0.05$ GB & $      73$ & $       0$ \\ 
ogm-mcfusion-WS-BASE* & $         0.05$ sec & $      6622.98$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-BASE-CGCF* & $         1.77$ sec & $      4460.12$ & $-\infty$ & $         0.01$ GB & $       3$ & $       0$ \\ 
ogm-mcfusion-WS-CGC* & $         1.44$ sec & $      4445.57$ & $-\infty$ & $         0.01$ GB & $      16$ & $       0$ \\ 
ogm-mcfusion-WS-CGC-CGCF* & $         1.75$ sec & $      4444.59$ & $-\infty$ & $         0.01$ GB & $      27$ & $       0$ \\ 
 ogm-mcfusion-WS-MC* & $         6.00$ sec & $      4444.58$ & $-\infty$ & $         0.03$ GB & $      22$ & $       0$ \\ 
ogm-mcfusion-WS-MC-CGCF* & $         6.31$ sec & $      4443.79$ & $-\infty$ & $         0.03$ GB & $      46$ & $       0$ \\ 
\bottomrule
\end{tabular}
\end{table}