### 2014-10-13 paper/tables/knott-3d-150.tex

\begin{table}[H]
\scriptsize
\centering
\caption{knott-3d-150 (8 instances)}
\label{tab:smalltable-knott-3d-150}
\begin{tabular}{lrrrrrr}
\toprule
           algorithm &       runtime     &         value &         bound &           mem &     best &      opt   \\ \midrule 
            ogm-CGC* & $         0.08$ sec & $     -4566.41$ & $     -4855.18$ & $         0.01$ GB & $       5$ & $       0$ \\ 
ogm-mcfusion-HC-BASE* & $         0.12$ sec & $         0.00$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE-CGCF* & $         0.16$ sec & $     -4160.88$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-CGC* & $         0.36$ sec & $     -4559.96$ & $-\infty$ & $         0.01$ GB & $       6$ & $       0$ \\ 
ogm-mcfusion-HC-CGC-CGCF* & $         0.36$ sec & $     -4565.61$ & $-\infty$ & $         0.01$ GB & $       6$ & $       0$ \\ 
 ogm-mcfusion-HC-MC* & $         0.88$ sec & $     -4559.96$ & $-\infty$ & $         0.03$ GB & $       6$ & $       0$ \\ 
ogm-mcfusion-HC-MC-CGCF* & $         0.88$ sec & $     -4565.61$ & $-\infty$ & $         0.03$ GB & $       6$ & $       0$ \\ 
ogm-mcfusion-WS-BASE* & $         0.03$ sec & $         0.00$ & $-\infty$ & $         0.01$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-BASE-CGCF* & $         0.12$ sec & $     -4565.95$ & $-\infty$ & $         0.01$ GB & $       5$ & $       0$ \\ 
ogm-mcfusion-WS-CGC* & $          NaN$ sec & $          NaN$ & $          NaN$ & $         0.01$ GB & $       4$ & $       0$ \\ 
ogm-mcfusion-WS-CGC-CGCF* & $         0.37$ sec & $     -4571.22$ & $-\infty$ & $         0.01$ GB & $       7$ & $       0$ \\ 
 ogm-mcfusion-WS-MC* & $         1.28$ sec & $     -4570.83$ & $-\infty$ & $         0.03$ GB & $       5$ & $       0$ \\ 
ogm-mcfusion-WS-MC-CGCF* & $         1.31$ sec & $     -4571.22$ & $-\infty$ & $         0.03$ GB & $       7$ & $       0$ \\ 
\bottomrule
\end{tabular}
\end{table}