### 2014-10-13 paper/tables/knott-3d-550.tex

\begin{table}[H]
\scriptsize
\centering
\caption{knott-3d-550 (8 instances)}
\label{tab:smalltable-knott-3d-550}
\begin{tabular}{lrrrrrr}
\toprule
           algorithm &       runtime     &         value &         bound &           mem &     best &      opt   \\ \midrule 
             ogm-ICM & $      2787.19$ sec & $   -126335.53$ & $-\infty$ & $         0.06$ GB & $       0$ & $       0$ \\ 
              ogm-KL & $       604.50$ sec & $   -127032.70$ & $-\infty$ & $         0.06$ GB & $       0$ & $       0$ \\ 
            ogm-LF-1 & $       987.28$ sec & $   -126356.35$ & $-\infty$ & $         0.07$ GB & $       0$ & $       0$ \\ 
\cmidrule{1-1} 
              MCR-CC & $     68016.21$ sec & $     -8187.14$ & $   -144703.64$ & $         1.20$ GB & $       0$ & $       0$ \\ 
           MCR-CCFDB & $     53527.56$ sec & $     -8187.14$ & $   -144703.64$ & $         0.47$ GB & $       0$ & $       0$ \\ 
       MCR-CCFDB-OWC & $     53386.77$ sec & $     -8187.14$ & $   -144703.64$ & $         0.47$ GB & $       0$ & $       0$ \\ 
\cmidrule{1-1} 
     MCI-CCFDB-CCIFD & $     53404.59$ sec & $     -8187.14$ & $   -144703.64$ & $         0.47$ GB & $       0$ & $       0$ \\ 
             MCI-CCI & $      3436.37$ sec & $   -101592.17$ & $   -136714.73$ & $         1.53$ GB & $       2$ & $       2$ \\ 
           MCI-CCIFD & $      3299.31$ sec & $   -120874.65$ & $   -136699.49$ & $         0.86$ GB & $       2$ & $       2$ \\ 
\cmidrule{1-1} 
            ogm-CGC* & $       595.15$ sec & $   -136188.55$ & $   -144703.64$ & $         0.11$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE* & $        17.17$ sec & $         0.00$ & $-\infty$ & $         0.11$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-BASE-CGCF* & $       118.24$ sec & $   -123617.74$ & $-\infty$ & $         0.19$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-CGC* & $       208.48$ sec & $   -136454.21$ & $-\infty$ & $         0.21$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-CGC-CGCF* & $       205.22$ sec & $   -136472.89$ & $-\infty$ & $         0.27$ GB & $       0$ & $       0$ \\ 
 ogm-mcfusion-HC-MC* & $       300.08$ sec & $   -136449.81$ & $-\infty$ & $         0.22$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-HC-MC-CGCF* & $       291.88$ sec & $   -136468.45$ & $-\infty$ & $         0.27$ GB & $       1$ & $       0$ \\ 
ogm-mcfusion-WS-BASE* & $         2.32$ sec & $         0.00$ & $-\infty$ & $         0.05$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-BASE-CGCF* & $       568.33$ sec & $   -136191.25$ & $-\infty$ & $         0.14$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-CGC* & $      1694.84$ sec & $   -136441.96$ & $-\infty$ & $         0.07$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-CGC-CGCF* & $      1900.61$ sec & $   -136491.63$ & $-\infty$ & $         0.15$ GB & $       0$ & $       0$ \\ 
 ogm-mcfusion-WS-MC* & $      3617.78$ sec & $   -136435.37$ & $-\infty$ & $         0.23$ GB & $       0$ & $       0$ \\ 
ogm-mcfusion-WS-MC-CGCF* & $      3773.77$ sec & $   -136494.37$ & $-\infty$ & $         0.28$ GB & $       4$ & $       0$ \\ 
\bottomrule
\end{tabular}
\end{table}