### 2014-11-05 paper/anytimefigure.tex
update experimental evaluation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pgfplotsset{every axis legend/.append style={
at={(1.0,1.2)},
anchor=north east}} 
\begin{figure*}[p]
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  restrict y to domain=0:4620,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false,
  ]
  %\addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data/knott-3d-150.data};
  %\addlegendentry{HC-CGC}

  \addplot[color=red,mark=x] table[x=time,y=CGC-planar]{data2/image-seg.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/image-seg.data};
  \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data2/image-seg.data};
 % \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/image-seg.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-MC]{data2/image-seg.data};
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/image-seg.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{image-seg}
  \end{subfigure}
  \begin{subfigure}[b]{0.33\textwidth}
    \centering
    \begin{tikzpicture}
    \begin{semilogxaxis}[
    restrict y to domain=-6000:-4400,
    xlabel = {runtime},
    xmin = 0,
    xmax = 4100,
    %ymin =  75282844.25,
    %ymax =  81774642.015,
    %xtick       = { 1 , 60 , 3600 },
    %xticklabels = { 1 sec. , 1 min.  , 1 hour },
    %ytick       = {75661636, 78718140, 81774643 },
    %yticklabels = {$75661636$, $78718140$, $81774643$ },
    width = 1.0\columnwidth,
    scaled ticks = false
    ]
    %\addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/knott-3d-150.data};
    %%\addlegendentry{HC-CGC}
    %
    \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/knott-3d-150.data};            
    \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/knott-3d-150.data};
    \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data2/knott-3d-150.data};  
    \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/knott-3d-150.data}; 
    \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/knott-3d-150.data};
    \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/knott-3d-150.data};

    \end{semilogxaxis}
    \end{tikzpicture}
    \caption{knott-3d-150}
  \end{subfigure}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  restrict y to domain=-40000:-24000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false,
  legend to name = named,
  legend columns=5
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/knott-3d-300.data};          \addlegendentry{HC-CGC}
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/knott-3d-300.data};               \addlegendentry{CGC}
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/knott-3d-300.data};       \addlegendentry{MC-I}
  \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data2/knott-3d-300.data};    \addlegendentry{Fusion-HC-MC}
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/knott-3d-300.data};    \addlegendentry{Fusion-HC-CGC}
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/knott-3d-300.data};  \addlegendentry{Fusion-WS-CGC}
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/knott-3d-300.data};          \addlegendentry{KL}

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{knott-3d-300}
  \end{subfigure}
  \newline
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  restrict y to domain=-80000:-60000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/knott-3d-450.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/knott-3d-450.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/knott-3d-450.data};
  \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data2/knott-3d-450.data};
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/knott-3d-450.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/knott-3d-450.data};
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/knott-3d-450.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{knott-3d-450}
  \end{subfigure}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  restrict y to domain=-10000000:-100000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/knott-3d-550.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/knott-3d-550.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/knott-3d-550.data};
  \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data2/knott-3d-550.data};
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/knott-3d-550.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/knott-3d-550.data};
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/knott-3d-550.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{knott-3d-550}
  \end{subfigure}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%\begin{subfigure}[H]
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  %restrict y to domain=-10000000:-100000,
  %restrict y to domain=-10000000:1200000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/seg-3d.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/seg-3d.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/seg-3d.data};
  \addplot[color=green,mark=o] table[x=time, y=DYNCC-HC-MC]{data2/seg-3d.data}; 
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/seg-3d.data}; 
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/seg-3d.data};
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/seg-3d.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{seg-3d}
  \end{subfigure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  %restrict y to domain=-10000000:-100000,
  %restrict y to domain=-10000000:100000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/socialnets.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/socialnets.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/socialnets.data};
  \addplot[color=green,mark=o] table[x=time, y=DYNCC-HC-MC]{data2/socialnets.data};
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/socialnets.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/socialnets.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{social nets}
  \end{subfigure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  %restrict y to domain=-10000000:-100000,
  %restrict y to domain=-10000000:100000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/normalizedsocialnets.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/normalizedsocialnets.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/normalizedsocialnets.data};
  \addplot[color=green,mark=o] table[x=time, y=DYNCC-HC-MC]{data2/normalizedsocialnets.data};
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/normalizedsocialnets.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data2/normalizedsocialnets.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{social nets}
  \end{subfigure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{subfigure}[b]{0.33\textwidth}
  \centering
  \begin{tikzpicture}
  \begin{semilogxaxis}[
  %restrict y to domain=-10000000:-100000,
  %restrict y to domain=-10000000:100000,
  xlabel = {runtime},
  xmin = 0,
  xmax = 4100,
  %ymin =  75282844.25,
  %ymax =  81774642.015,
  %xtick       = { 1 , 60 , 3600 },
  %xticklabels = { 1 sec. , 1 min.  , 1 hour },
  %ytick       = {75661636, 78718140, 81774643 },
  %yticklabels = {$75661636$, $78718140$, $81774643$ },
  width = 1.0\columnwidth,
  scaled ticks = false
  ]
  \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data2/modularity-clustering.data};
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data2/modularity-clustering.data};
  \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data2/modularity-clustering.data};
  \addplot[color=green,mark=o] table[x=time, y=DYNCC-HC-MC]{data2/modularity-clustering.data};
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/modularity-clustering.data};
  \addplot[color=orange,mark=x] table[x=time, y=DYNCC-HC-CGC]{data2/modularity-clustering.data};
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data2/modularity-clustering.data};

  \end{semilogxaxis}
  \end{tikzpicture}
  \caption{modularity clustering}
  \end{subfigure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \begin{center}
  \hypersetup{linkcolor = black}
  \ref{named}
  \hypersetup{linkcolor = red}
  \end{center}
  \caption{
    Among all tested solvers, the proposed method is the only
    one scaling to large 3D data.
    Except for the small 3D block, CC-Fusion finds very low
    energies one to three magnitudes faster 
    than any other method.
    With increasing block size, is outperformed even more.
    For the 2D datasets CGC is very fast, but our proposed
    solver finds lower energies while still being faster
    than the global optimal multicut solver.
  }
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%