### 2014-11-10 paper/inputs/fig_fusion.tex
mystuff

%!TEX root = ../egpaper_for_review.tex

\tikzstyle{nS}=[circle,minimum size = 0.6cm,inner sep = 0pt,draw, font=\small,align=left]
\tikzstyle{nD}=[shape = circle split,rotate=45,minimum size = 0.6cm,inner sep = 0pt,draw, font=\small,align=left]

\tikzstyle{eS}=[thick,minimum size = 0.5cm,inner sep = 0pt,draw, font=\small,align=left]
\tikzstyle{eC}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=red, font=\small,align=left]
\tikzstyle{eNC}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=green!70!black, font=\small,align=left, line width=1mm]
\tikzstyle{eNS}=[fill=white,minimum size = 0.5cm,inner sep = 0pt, font=\tiny,align=left]
\tikzstyle{eCC}=[very thick,dotted,minimum size = 0.5cm,inner sep = 0pt,draw=black, font=\small,align=left, draw=red!70!black, line width=1mm]

\tikzset{
  head/.style = {fill = orange!90!blue,
                 label = center:\textsf{\Large H}},
  tail/.style = {fill = blue!70!yellow, text = black,
                 label = center:\textsf{\Large T}}
}

\begin{center}
\begin{figure}[ht]
\begin{tiny}

\begin{subfigure}[t]{0.7\linewidth}
    \resizebox{\linewidth}{!}{
        \begin{tikzpicture}[scale=1]%[x=1 cm, y=1 cm]

                \draw (2,2-2) node[nS,fill=red!60] (n0) {$0$}; 
                \draw (4,2-2) node[nS,fill=red!60] (n1) {$1$};
                \draw (2,0-2) node[nS,fill=green!60] (n3) {$2$}; 
                \draw (4,0-2) node[nS,fill=green!60] (n4) {$3$};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                % 1. FUSION
                \draw (2+4,2) node[nS,fill=red!60] (n0) {$0$}; 
                \draw (4+4,2) node[nS,fill=green!60] (n1) {$1$};
                \draw (2+4,0) node[nS,fill=red!60] (n3) {$2$}; 
                \draw (4+4,0) node[nS,fill=green!60] (n4) {$3$};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                \draw (2+4,2-4) node[nD, circle split part fill={red!60, red!60}] (n0) {$ $}; 
                \draw (4+4,2-4) node[nD, circle split part fill={red!60, green!60}] (n1) {$ $};
                \draw (2+4,0-4) node[nD, circle split part fill={green!60, red!60}] (n3) {$ $}; 
                \draw (4+4,0-4) node[nD, circle split part fill={green!60, green!60}] (n4) {$ $};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                % 2. FUSION
                \draw (2+8,2) node[nS,fill=blue!60] (n0) {$0$}; 
                \draw (4+8,2) node[nS,fill=green!60] (n1) {$1$};
                \draw (2+8,0) node[nS,fill=blue!60] (n3) {$2$}; 
                \draw (4+8,0) node[nS,fill=green!60] (n4) {$3$};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                \draw (2+8,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
                \draw (4+8,2-4) node[nD, circle split part fill={red!60, green!60}] (n1) {$ $};
                \draw (2+8,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
                \draw (4+8,0-4) node[nD, circle split part fill={green!60, green!60}] (n4) {$ $};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                % 3. FUSION
                \draw (2+12,2) node[nS,fill=red!60] (n0) {$0$}; 
                \draw (4+12,2) node[nS,fill=blue!60] (n1) {$1$};
                \draw (2+12,0) node[nS,fill=red!60] (n3) {$2$}; 
                \draw (4+12,0) node[nS,fill=blue!60] (n4) {$3$};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                \draw (2+12,2-4) node[nD, circle split part fill={red!60, red!60}] (n0) {$ $}; 
                \draw (4+12,2-4) node[nD, circle split part fill={red!60, blue!60}] (n1) {$ $};
                \draw (2+12,0-4) node[nD, circle split part fill={green!60, red!60}] (n3) {$ $}; 
                \draw (4+12,0-4) node[nD, circle split part fill={green!60, green!60}] (n4) {$ $};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                % 4. FUSION
                \draw (2+16,2) node[nS,fill=magenta!60] (n0) {$0$}; 
                \draw (4+16,2) node[nS,fill=blue!60] (n1) {$1$};
                \draw (2+16,0) node[nS,fill=magenta!60] (n3) {$2$}; 
                \draw (4+16,0) node[nS,fill=blue!60] (n4) {$3$};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

                \draw (2+16,2-4) node[nD, circle split part fill={red!60, magenta!60}] (n0) {$ $}; 
                \draw (4+16,2-4) node[nD, circle split part fill={red!60, blue!60}] (n1) {$ $};
                \draw (2+16,0-4) node[nD, circle split part fill={green!60, magenta!60}] (n3) {$ $}; 
                \draw (4+16,0-4) node[nD, circle split part fill={green!60, blue!60}] (n4) {$ $};
                \path
                    (n0) edge[eS] node[eNS]{$w_{01}$} (n1) 
                    (n3) edge[eS] node[eNS]{$w_{23}$} (n4) 
                    (n0) edge[eS] node[eNS]{$w_{02}$} (n3)
                    (n1) edge[eS] node[eNS]{$w_{13}$} (n4)
                ;

        \end{tikzpicture}
    }
\caption{b}
\label{fig:f_b}
\end{subfigure}

\end{tiny}
\caption{
w
}\label{fig:fusion_abiguty}
\end{figure}
\end{center}