### 2014-11-11 paper/inputs/fig_fusion.tex
made fusion figure smaller

%!TEX root = ../egpaper_for_review.tex

\tikzstyle{nS}=[circle,minimum size = 0.6cm,inner sep = 0pt,draw, font=\small,align=left]
\tikzstyle{nD}=[shape = circle split,rotate=45,minimum size = 0.6cm,inner sep = 0pt,draw, font=\small,align=left]

\tikzstyle{eS}=[thick,minimum size = 0.5cm,inner sep = 0pt,draw, font=\small,align=left]
\tikzstyle{eC}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=red, font=\small,align=left]
\tikzstyle{eNC}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=green!70!black, font=\small,align=left, line width=1mm]
\tikzstyle{eNS}=[fill=white,minimum size = 0.5cm,inner sep = 0pt, font=\tiny,align=left]
\tikzstyle{eCC}=[very thick,dotted,minimum size = 0.5cm,inner sep = 0pt,draw=black, font=\small,align=left, draw=red!70!black, line width=1mm]

\tikzset{
  head/.style = {fill = orange!90!blue,
                 label = center:\textsf{\Large H}},
  tail/.style = {fill = blue!70!yellow, text = black,
                 label = center:\textsf{\Large T}}
}

\begin{center}
\begin{figure}[ht]
\resizebox{\linewidth}{!}{
    \begin{tikzpicture}[scale=1]%[x=1 cm, y=1 cm]

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3.5) -- (1-2,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\Large{proposal edge} \\ \Large{labeling \quad $\bm{x'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (1-2,3-3.5) -- (1-2,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{proposal edge} \\ \Large{labeling \quad $\bm{x''}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3.5) -- (5,6.5) node [black,midway,xshift=-1.2cm,rotate=90,align=left] 
            {\Large{proposal node} \\ \Large{labeling \quad $\bm{l'}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,3-3.5) -- (5,6.5-4) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{proposal node} \\ \Large{labeling \quad $\bm{l''}$}};

            \draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt]
            (5,-4.5) -- (5,-1.5) node [black,midway,xshift=-1.2cm, rotate=90,align=left] 
            {\Large{Binary fusion} \\ \Large{move problem}};

            \draw (2-2,2+4) node[nS] (n0) {$0$}; 
            \draw (4-2,2+4) node[nS] (n1) {$1$};
            \draw (2-2,0+4) node[nS] (n3) {$2$}; 
            \draw (4-2,0+4) node[nS] (n4) {$3$};
            \path
                (n0) edge[eNotCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eNotCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eCut] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2-2,2) node[nS] (n0) {$0$}; 
            \draw (4-2,2) node[nS] (n1) {$1$};
            \draw (2-2,0) node[nS] (n3) {$2$}; 
            \draw (4-2,0) node[nS] (n4) {$3$};
            \path
                (n0) edge[eCut] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eCut] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eNotCut] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eNotCut] node[eNS]{$w_{13}$} (n4)
            ;

            % first segmentation as coloring
            \draw (2+3.5,2+4) node[nS,fill=red!60] (n0) {$0$}; 
            \draw (4+3.5,2+4) node[nS,fill=red!60] (n1) {$1$};
            \draw (2+3.5,0+4) node[nS,fill=green!60] (n3) {$2$}; 
            \draw (4+3.5,0+4) node[nS,fill=green!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 1. FUSION
            \draw (2+3.5,2) node[nS,fill=red!60] (n0) {$0$}; 
            \draw (4+3.5,2) node[nS,fill=green!60] (n1) {$1$};
            \draw (2+3.5,0) node[nS,fill=red!60] (n3) {$2$}; 
            \draw (4+3.5,0) node[nS,fill=green!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+3.5,2-4) node[nD, circle split part fill={red!60, red!60}] (n0) {$ $}; 
            \draw (4+3.5,2-4) node[nD, circle split part fill={red!60, green!60}] (n1) {$ $};
            \draw (2+3.5,0-4) node[nD, circle split part fill={green!60, red!60}] (n3) {$ $}; 
            \draw (4+3.5,0-4) node[nD, circle split part fill={green!60, green!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 2. FUSION
            \draw (2+7,2) node[nS,fill=red!60] (n0) {$0$}; 
            \draw (4+7,2) node[nS,fill=blue!60] (n1) {$1$};
            \draw (2+7,0) node[nS,fill=red!60] (n3) {$2$}; 
            \draw (4+7,0) node[nS,fill=blue!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+7,2-4) node[nD, circle split part fill={red!60, red!60}] (n0) {$ $}; 
            \draw (4+7,2-4) node[nD, circle split part fill={red!60, blue!60}] (n1) {$ $};
            \draw (2+7,0-4) node[nD, circle split part fill={green!60, red!60}] (n3) {$ $}; 
            \draw (4+7,0-4) node[nD, circle split part fill={green!60, blue!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 3. FUSION
            \draw (2+10.5,2) node[nS,fill=green!60] (n0) {$0$}; 
            \draw (4+10.5,2) node[nS,fill=red!60] (n1) {$1$};
            \draw (2+10.5,0) node[nS,fill=green!60] (n3) {$2$}; 
            \draw (4+10.5,0) node[nS,fill=red!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+10.5,2-4) node[nD, circle split part fill={red!60, green!60}] (n0) {$ $}; 
            \draw (4+10.5,2-4) node[nD, circle split part fill={red!60, red!60}] (n1) {$ $};
            \draw (2+10.5,0-4) node[nD, circle split part fill={green!60, green!60}] (n3) {$ $}; 
            \draw (4+10.5,0-4) node[nD, circle split part fill={green!60, red!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 4. FUSION
            \draw (2+14,2) node[nS,fill=green!60] (n0) {$0$}; 
            \draw (4+14,2) node[nS,fill=blue!60] (n1) {$1$};
            \draw (2+14,0) node[nS,fill=green!60] (n3) {$2$}; 
            \draw (4+14,0) node[nS,fill=blue!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+14,2-4) node[nD, circle split part fill={red!60, green!60}] (n0) {$ $}; 
            \draw (4+14,2-4) node[nD, circle split part fill={red!60, blue!60}] (n1) {$ $};
            \draw (2+14,0-4) node[nD, circle split part fill={green!60, green!60}] (n3) {$ $}; 
            \draw (4+14,0-4) node[nD, circle split part fill={green!60, blue!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 5. FUSION
            \draw (2+17.5,2) node[nS,fill=blue!60] (n0) {$0$}; 
            \draw (4+17.5,2) node[nS,fill=red!60] (n1) {$1$};
            \draw (2+17.5,0) node[nS,fill=blue!60] (n3) {$2$}; 
            \draw (4+17.5,0) node[nS,fill=red!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+17.5,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
            \draw (4+17.5,2-4) node[nD, circle split part fill={red!60, red!60}] (n1) {$ $};
            \draw (2+17.5,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
            \draw (4+17.5,0-4) node[nD, circle split part fill={green!60, red!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 6. FUSION
            \draw (2+21,2) node[nS,fill=blue!60] (n0) {$0$}; 
            \draw (4+21,2) node[nS,fill=green!60] (n1) {$1$};
            \draw (2+21,0) node[nS,fill=blue!60] (n3) {$2$}; 
            \draw (4+21,0) node[nS,fill=green!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+21,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
            \draw (4+21,2-4) node[nD, circle split part fill={red!60, green!60}] (n1) {$ $};
            \draw (2+21,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
            \draw (4+21,0-4) node[nD, circle split part fill={green!60, green!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            % 7. FUSION
            \draw (2+24.5,2) node[nS,fill=blue!60] (n0) {$0$}; 
            \draw (4+24.5,2) node[nS,fill=yellow!60] (n1) {$1$};
            \draw (2+24.5,0) node[nS,fill=blue!60] (n3) {$2$}; 
            \draw (4+24.5,0) node[nS,fill=yellow!60] (n4) {$3$};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw (2+24.5,2-4) node[nD, circle split part fill={red!60, blue!60}] (n0) {$ $}; 
            \draw (4+24.5,2-4) node[nD, circle split part fill={red!60, yellow!60}] (n1) {$ $};
            \draw (2+24.5,0-4) node[nD, circle split part fill={green!60, blue!60}] (n3) {$ $}; 
            \draw (4+24.5,0-4) node[nD, circle split part fill={green!60, yellow!60}] (n4) {$ $};
            \path
                (n0) edge[eBlack] node[eNS]{$w_{01}$} (n1) 
                (n3) edge[eBlack] node[eNS]{$w_{23}$} (n4) 
                (n0) edge[eBlack] node[eNS]{$w_{02}$} (n3)
                (n1) edge[eBlack] node[eNS]{$w_{13}$} (n4)
            ;

            \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0.8cm]
            (5,-6) -- (29,-6) node [black,midway,xshift=0cm,yshift=-1cm] 
            {
                \Large{For all different colorings $\bm{l''}$, the binary fusion move subproblem is different.}
            };

            \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0.8cm]
            (5,2) -- (29,2) node [black,midway,xshift=0cm,yshift=1cm] 
            {
                \Large{All these node labelings encode the same partition}
            };

            \draw[black] (5,-1) -- (29,-1);

    \end{tikzpicture}
}
\caption{
todo: write me (why normal fusion does not work)
}\label{fig:fusion_abiguty}
\end{figure}
\end{center}