### 2014-11-13 paper/evaluation.tex

\subsection{Evaluation}
For the evaluation of the diffrent methods on the datasets introduced in Sec.~\ref{sec:datasets},
we show zoomed anytime plots in  Fig.~\ref{fig:anytime} and variation of information (VOI)~\cite{meila-2003}
and rand index (RI)~\cite{rand-1971} of the final solutions in Tab.~\ref{tab:eval}.
Anytime plots with no zooming and a detailed evalution are given in supplementray material.

For \emph{social-nets} CC-Fusion methods provides for the first minutes the best results.
Only CGC and HC-CGC are able to find better solutions after more then 1000 seconds.
MC-I and MC-R can even not be applied on such large problems.
We belief that with better proposal generators, which are more suited for such network problems,
we can improve CCFusion. One candidate for such a generated would be a scalable implementations of the 
PIVOT algorithm.

For \emph{modularity-clustering} CC-Fusion performs on par to competitive methods, even it 
and the used parameters was not designed and chosen for this type of problem.
Especially is does a better job than MC-I.

For \emph{image-seg} CC-Fusion is faster than other methods and competitive in terms of energy, VOI, and RI.
Because, the models are designed to have a high boundary recall (oversegmentation), 
classical fusion, which return undersegmentations, with best VOI but worse RI and energy.
Proposals generated by EHC seemed to be a bit better than WS-based ones.

For the \emph{knott-datasets} similar holds as for image-seg. 
With increasing problem size CC-Fusion-HC-MC and CC-Fusion-WS-MC have a better performance compared to competitive methods, \cf Fig.~\ref{fig:anytime}(b-e).
Also in terms of VOI and RI they are only slightly worse than the global optimal solution found by MC-I.
The initialization of CGC by HC denoted by HC-CGC also improves the performance compared to native CGC.
For the largest 3D volume seg-3d, HC-CGC gives the first usefull solution, \cf  Fig.~\ref{fig:anytime}(f).
However, after a view minutes  CC-Fusion-HC-MC and CC-Fusion-WS-MC give much better results and also overall best in terms of energy, VOI, and RI.
%
Pure EHC, Fusion and PIVOT-BOEM do not give useful results on any dataset.