### 2014-11-13 paper/abstract-new.tex
add new abstract

Correlation clustering, or multicut partitioning,
is a widely used in image segmentation for
partitioning an undirected graph or image with positive and negative edge weights 
such that the sum of cut edge weights is minimized.
%
Due to its NP-hardness, exact solvers do not scale and approximative solvers often give nonsatisfying results.

We investigate scalable methods for correlation clustering.
To this end we define fusion moves for the correlation clustering problem.
Our algorithm iteratively fuse the current and a proposed partitioning which  monotonously improve
the partitioning and maintain a valid partitioning at all times.
Furthermore, it scales to larger datasets, gives near optimal solutions, and at the same time shows
a very good anytime performance.