### 2014-11-14 paper/abstract-new.tex
typos found by paul

Correlation clustering, or multicut partitioning,
is widely used in image segmentation for
partitioning an undirected graph or image with positive and negative edge weights 
such that the sum of cut edge weights is minimized.
%
Due to its NP-hardness, exact solvers do not scale and approximative solvers often give non satisfying results.

We investigate scalable methods for correlation clustering.
To this end we define fusion moves for the correlation clustering problem.
Our algorithm iteratively fuses the current and a proposed partitioning which  monotonously improves
the partitioning and maintains a valid partitioning at all times.
Furthermore, it scales to larger datasets, gives near optimal solutions, and at the same time shows
a good anytime performance.