### 2015-01-19 rebuttal.txt
add rebuttal

We thank all reviews for their constructive comments.

All reviewers mention the
 * clear presentation
 * Detailed experiments
 * improves state-of-the-art

---
**R1: Analogues of the proposed EHC in the literature?**
???

**R1: Misses important prior work outside computer vision (multi-level graph partitioning algorithm)**
As stated by Buluc et al it is assumed that edge weights are positive and a volume constrained are used. 
The METIS software can not applied on our graphs with include positive and _negative_ weighted edges.

From the algorithmic view our proposed method is different to multi-level graph partitioning. 
We refine the graph with respect to the current labeling and solve the problem on the finer graph and not on a coarser (In Fig 5 (c) is finer than (a)).

**Remarks:**
We will improve the readability of the figures  
---

**R2: The proposed method is rather straightforward**
We belief that this is a clear pro of our method and not a con. Important is its novelty.

**R2: The proposed fusion move's convergence is sometimes worse than the CGC**
For the social nets we belief that our proposal generator is not a good choice and parameters can be tuned. But even whit this default it works fine.

**R2: Proposal generation algorithms themselves are not validated**
Due to lack of space we skipped this. However, (E)HC is compared in Fig. 9. One can see that proposals are not globally meaningful.

**R2: The effect of the proposed third criterion "size" **
See Fig. 8. If the size is to large the subproblems become intractable.

**R2: Generated proposals not divergent**
Empirically we observe that the proposals are divergent. 
Of cause that requires that the signal to noise ratio is not to small.

**R2: parameter selection for the proposal generation is verified with only a single instance**
We never claim that we have found the optimal parameter. We expect that we can improve the results by tuning the parameters what we have _not_ done.

**R2: Theorem 1 and eq. 5**
R2 is right it is "for any y" and "\breve{y}=y". We will make this more clear.
While this looks obvious on first glance it isn't on the second. 
We feel that its important to proof this.

**R2: no theoretical guarantee**
Theorem 1 and Sec 4.2 give some theoretical interpretation, e.g. that the method is not energy increasing.
Since there is no free lunch we have to give up the lower bound for faster speed.
---

**R3: Will the implementation become available if paper is accepted?**
We will our implementation public available if paper is accepted. 
Beyond the theoretical contribution in includes several implementation issues which we could not discuss to lack of space.